Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-08-09 paper/ms.tex
dicussion section to 90%

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs (7 subfields), this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs (3 subfields). The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs (1 subfield),
and the field around the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs (1 subfield).

@@

Figure \ref{fig:correlation_parameters} shows the main results in a
$\theta_{0}-\beta$  plane where the average and standard deviation
over the mocks is shown in comparison with the result derived from
observations.  The error bars in this Figure represents the standard
deviation of the ACF over all the sub-fields in the 15 mock
observations. These error bars are larger than the statistical
uncertainty from the fitting procedure on a single field.

In the left panel Figure \ref{figure:correlation_parameters} we see
how the observational ACF measured by \cite{Hayashino2004} (green dot)
is successful in reducing the total number of possible models. Only
those with angular-correlation length within
$15^{\prime\prime}<\theta_{0}<23^{\prime\prime}$ are considered to
reproduce observations.     {\bf Cual es el resultsdo de Hayashino
  para $\theta_0$?}

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the preferred
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method. To make this selection we consider a that
a model is consistent with observations if there is a $1-\sigma$
overlap between both the correlation length $\theta_0$ and the power
$\beta$ in the correlation function. Comparing this Figure with the
results shown in Figure \ref{fig:restriction_mock} we see that the
models that were removed were those with highest minimal masses
$M_{\rm min}>10^{11.2}$, for all the values of $M_{\rm max}$, which
had preferentially high occupation fractions of $f_{\rm occ}> 0.5$.

@@

Matching the galaxy surface number density statistics to
our mock  data sets the median
mass of all successful models in the range $(10^{10}-10^{12})$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 9000 possible combination of parameters.

We find that to make a physical interpretation of these models it is
useful to do it in terms of the size of halo mass range. We now use the
variable $\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm  min}$, together
with the occupation fraction $f_{\rm occ}$ to help us build a classification
of all the successful models into three families:

@@

At the low mass end, such cut can be readily interpreted in terms of the
minimal halo star formation rate needed to produce the necessary \ly 
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increasing with halo
mass, the cut at higher halo masses requires a different
explanation. There are two different physical scenarios to explain
this.

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems which, due to the resonant nature of the
\ly line that increase the probability of absortion  of\ly photons, produce
high absorption of \ly photons but not of continuum or
other n-resonant lines \citep{Laursen2009,ForeroRomero2011}.

A second scenario is that larger systems have more extended gaseous
envelopes which, due to resonance effects of the \ly line, leading to
a low surface brightness and a broader line, making these systems less
observable in narrow band filter surveys \citep{Zheng2010}.

The preference for narrow $\Delta M$ ranges to hosting LAEs together
with the minor presence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun \  shows that our models support theoretical and observational
insights where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012}, \citep{Shapley2003}

\subsection{Comparison against results from blind surveys}

@@

We also find interesting that our ACF analysis is also not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models can only be
considered unfavorable based on the density and abundance observations
as we have described in the previous subsection.

@@

Considering the additional evidence for a low escape fraction and 
assuming a direct relation between $f_{esc}$ and $f_{occ}$ we can
say that the preferred models are in families (1) and (2) with a
clear majority composed by those with narrow mass range.

Abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} work based on observational
results for Lyman Break Galaxies. In those studies the minimum halo
mass considered to be relevant in their analysis is
$10^{11.4}$\hMsun. For that mass, \cite{Behroozi2013a} report a
stellar mass mass around $(1.0\pm0.3)\times 10^{9.0}$
\hMsun, while their star formation rate is in the range $0.6\pm 0.2$
Msun yr$^{-1}$, which nevertheless are close to the lower bound of
values inferred for LAEs at high redshift
\citep{Gawiser2007,Nilsson2009,Pentericci2009}.

In our results, all the preferred models have a halo mass range lower
than the minimum of $M_{\rm min}<10^{11.4}$\hMsun considerend in
abundance matching at $z=3$. This mass scale is also strictly superior to
the great majority of  $M_{\rm max}$ values allowed in the models with
low escape fraction.   This suggests that a detailed and careful study
of the spectral and photometric properties of LAEs coupled to the kind
of analysis performed in this paper can be a guide in the study of the
properties of low mass dark matter halos at $z=3.1$, extending the
capabilities of abundance matching methods.

@@

... All the software to produce the results in this paper is publicly
available.

... The raw catalogs can be obtained from the MultiDark database but
can also be obtained in the repository of this paper on github.

>>>>

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs (7 subfields), this
field has been known to harbor a region with a large density excess of
galaxies. The second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a
total of $0.58$ deg$^2$ and $386$ LAEs (3 subfields). The third and
fourth fields are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and
$196$ LAEs (1 subfield), and the field around the Great Observatory
Optical Deep Survey  (GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs (1
subfield).

@@

Figure \ref{fig:correlation_parameters} shows the main results in a
$\theta_{0}-\beta$  plane where the average and standard deviation
over the mocks is shown in comparison with the result derived from
observations.  The error bars in this Figure represents the standard
deviation of the ACF over all the 15 mock fields used to make the
computation.  These error bars are larger than the statistical
uncertainty from the fitting procedure on a single field.

In the left panel Figure \ref{figure:correlation_parameters} we see
how the observational ACF measured by \cite{Hayashino2004} (green dot)
is successful in reducing the total number of possible models. Only
those with angular-correlation length within
$15^{\prime\prime}<\theta_{0}<23^{\prime\prime}$ are considered to
reproduce observations. {\bf Cual es el resultsdo de Ouchi
  para $\theta_0$?}

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the preferred
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method. To make this selection we consider that
a model is consistent with observations if there is a $1-\sigma$
overlap between both the correlation length $\theta_0$ and the power
$\beta$ . Comparing this Figure with the results shown in Figure
\ref{fig:restriction_mock} we see that the models that were removed
were those with highest minimal masses $M_{\rm min}>10^{11.2}$, for
all the values of $M_{\rm max}$, with preferentially high
occupation fractions of $f_{\rm occ}> 0.5$.

@@

In the Results Section we start by matching the rough numbers for the
galaxy surface number density. This sets the median mass of all
successful models to the broad range $(10^{10}-10^{12})$ \hMsun for
$M_{\rm min}$ and $M_{\rm max}$. We include additional criteria on the
number of mock surveys that must be consistent observations
and the information from the angular correlation function, reducing
the number of feasible models. We end up with $\sim 50$ models out of
the initial $9000$ possible combination of parameters.

In order to make a physical interpretation of these models we use 
the halo mass range, defined as $\Delta M=\log_{10}M_{\rm max} -
\log_{10}M_{\rm  min}$, together with the occupation fraction $f_{\rm
  occ}$ to classify all the successful models into three families:

@@

At the low mass end, such cut can be readily interpreted in terms of the
minimal halo star formation rate needed to produce the minimal \ly 
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increasing with halo
mass, the cut at higher halo masses requires a different
justification. There are two complementary physical scenarios that can
provide an explanation.

One sceneario can be presented in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems. Due to the resonant nature of the
\ly line the probability of absortion  of \ly photons increases in
massive systems, producing high absorption of \ly photons but not of
continuum or other non-resonant lines
\citep{Laursen2009,ForeroRomero2011}.

A second scenario is that larger systems have more extended gaseous
envelopes which. Again, due to resonance effects of the \ly line,
these systems can present a low surface brightness and a broader line,
making these systems less observable in narrow band filter surveys
\citep{Laursen2009,Zheng2010}.

The preference for narrow $\Delta M$ ranges to host LAEs 
and the minoritary presence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun follows theoretical and observational
insights where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012,Shapley2003}.

\subsection{Comparison against results from galaxy surveys}

@@

We also find that an ACF analysis on a single dataset is not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models cannot be
considered unfavorable based on a clustering analysis only.

@@

Abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} work based on observational
results for Lyman Break Galaxies (LBG). In those studies the minimum halo
mass considered to be relevant in their analysis is
$10^{11.4}$\hMsun. For that mass, \cite{Behroozi2013a} report a
stellar mass mass around $(1.0\pm0.3)\times 10^{9.0}$
\hMsun, while their star formation rate is in the range $0.6\pm 0.2$
Msun yr$^{-1}$, which nevertheless are close to the lower bound of
values inferred for LAEs at high redshift
\citep{Gawiser2007,Nilsson2009,Pentericci2009}.

In our results, all the preferred models have a halo mass range lower
than the minimum of $M_{\rm min}<10^{11.4}$\hMsun considerend in
abundance matching at $z=3$. This shows that our results also confirm
the expectations of LAEs to be hosted by less massive halos than LBGs.
A detaile analysis of the spectral and photometric properties of LAEs
coupled to the kind of analysis performed in this paper can be a guide
in the study of the properties of low mass dark matter halos at
$z=3.1$, extending the capabilities of abundance matching methods.

\subsection{Possible caveats to our method}

... Single LAE in a halo.

... Mass definition for different halo finders.

@@

All the software, raw data and processed data to produce the results
and plots in this paper are publicly available in a github
repository \footnote{https://github.com/forero/CosmicVarianceLAES}. Most
of the code to produce the plots can be found as an Ipython notebook
\citep{IPython} in the same repository.
